---
title: 'Linear Regression Approach'
date: "May. 1, 2023"
output:
  pdf_document: default
  html_document: default
---
# install and load packages
```{r}
#install.packages("lars")
#install.packages("ggplot2")
#install.packages("Amelia")
#install.packages("caTools")
#install.packages("caret")
#install.packages("forecast")
#install.packages("agricolae")
#install.packages("Metrics")
library(data.table)
library(readr)
library(tidyverse)
library(Amelia)
library(caTools)
library(corrplot)
library(caret)
library(forecast)
library(ggplot2)
library(agricolae)
library(faraway)
library(MASS)
library(lars)
library(caret)
library(Metrics)
```

## Import dataset

```{r}
bike <-read_csv("C:\\Users\\willifornia\\Documents\\R_stat429\\Group4BikeDataAutumn_Revised.csv")

```

## Factorize non-numerical variables
```{r}

bike$Holiday <- as.factor(bike$Holiday)
bike$Weekday <- as.factor(bike$Weekday)
summary(bike)
```

## Produce correlation matrix

```{r}
num.col <- sapply(bike, is.numeric)
cor <- cor(bike[,num.col])
corrplot(cor, addCoef.col =1, tl.cex = 1, cl.cex = 1)

```
## Bike vs. Hours & Temp
```{r}
ggplot(bike, aes(Hour, Bike))+
  geom_point(position = position_jitter(w=1,h=0), aes(color = Temperature), size = 3)+
  scale_color_gradientn(colours = c('darkblue', 'lightblue','yellow', 'darkorange', 'darkred')) +
  theme(panel.background = element_rect(fill = "light grey"),
        axis.title = element_text(size = 20))
```

## Bike vs. Humidity
```{r}
ggplot(bike, aes(Humidity, Bike)) +
  geom_point(aes(col = Humidity), alpha = 0.5,size=3) +
  scale_y_continuous(n.breaks = 10) +
  scale_color_gradientn(colours = c('red', 'purple', 'blue', 'darkblue')) +
  theme(panel.background = element_rect(fill = "gray90"),
        axis.title = element_text(size = 20))
```
## Bike vs. Windspeed
```{r}
ggplot(bike, aes(Wind_Speed,Bike))+
  geom_point(aes(col = Wind_Speed), alpha = 0.5,size=3)+
  scale_y_continuous(n.breaks =10)+
  scale_color_gradientn(colours = c('blue', 'darkblue', 'purple','red'))+
  theme(panel.background = element_rect(fill = "gray90"),
        axis.title = element_text(size = 20))
```
## Bike vs. Visibility
```{r}
ggplot(bike, aes(Visibility,Bike))+
  geom_point(aes(col = Visibility), alpha = 0.5,size=3)+
  scale_y_continuous(n.breaks =10)+
  scale_color_gradientn(colours = c('red', 'pink','purple', 'blue', 'darkblue'))+
  theme(panel.background = element_rect(fill = "gray90"),
        axis.title = element_text(size = 20))
```
## Bike vs.Weekday
```{r}
ggplot(bike, aes(x = Weekday, y = Bike, fill = Weekday)) +
  geom_bar(stat = "identity") +
  labs(x = "Weekday", y = "Sum of bikes", title = "Sum of bikes by weekday") +
  scale_fill_manual(values = c("Monday" = "red", "Tuesday" = "orange", "Wednesday" = "yellow", "Thursday" = "green", "Friday" = "blue", "Saturday" = "brown", "Sunday" = "violet"))
```
## Weekday Difference
```{r}
mean_bikes <- aggregate(Bike ~ Weekday, data = bike, FUN = mean)
ggplot(mean_bikes, aes(x = Weekday, y = Bike)) +
  geom_bar(stat = "identity") +
  labs(x = "Weekday", y = "Average number of bikes rented")
```

### Linear Regression Analysis
## Remove "Functioning_Day" column
```{r}
bike <- subset(bike,select=-c(Functioning_Day))
```

## Split data into train and test set (75% train, 25% test)
```{r}
set.seed(123)
train_idx<-createDataPartition(bike$Bike,p=0.75,list=FALSE)
traindata <-bike[train_idx,]
testdata <-bike[-train_idx,]
```
## Examine the collinearity presented in the regression model
```{r}
bikemod_vif <-lm(Bike~.-Date,data=traindata)
vif(bikemod_vif)
```
## Remove Dew_point_temperature to address collinearity problem

```{r}
bikemod_simple <- lm(Bike~.-Date-Dew_point_temperature,data=traindata)
summary(bikemod_simple)
vif(bikemod_simple) #VIF are good after removing Dew_point_temperature
```
## Accuracy of the linear regression model
```{r}
predss <- predict(bikemod_simple,newdata = testdata)
print(mae(predss,testdata$Bike)) #Calculate MAE
```
## produce diagnostic plots for the linear regression model
par(mfrow=c(2,2))
plot(bikemod_simple,add.smooth = FALSE) 


## Use Box-Cox function to find optimal power transformation for response variable
```{r}
bc<-boxcox(bikemod_simple)
(lambda_trans <- bc$x[which.max(bc$y)]) #Choose 0.5 as our power transformation for bike usage
bikemod_boxcox<-lm(Bike^0.5~.-Date-Dew_point_temperature,data=bike)
summary(bikemod_boxcox)
par(mfrow=c(2,2))
plot(bikemod_boxcox,add.smooth = FALSE) 
```

## An illustration for adding polynomial terms of the Hour variable (power of 8)
```{r}
ggplot(traindata, aes(Hour, Bike^0.5))+
  geom_point(position = position_jitter(w=1,h=0), aes(color = Temperature),size=3)+
  scale_color_gradientn(colours = c('darkblue', 'lightblue','yellow', 'darkorange', 'darkred'))+
  stat_smooth(formula = y~poly(x,8, raw = TRUE), lty = 'dotted')+
  theme(panel.background = element_rect(fill = "gray90"),
        axis.title = element_text(size = 20))
```
## Add polynomial and interaction terms in the model
```{r}
bikemod_transform <-lm(Bike^0.5~.-Date-Dew_point_temperature+
                         poly(Hour,8,raw=T)+poly(Humidity,4,raw=T)+
                         Humidity*Temperature+
                         Solar_Radiation*Humidity+
                         Solar_Radiation*Temperature,data=traindata)
par(mfrow=c(2,2))
plot(bikemod_transform,add.smooth = FALSE)
summary(bikemod_transform)
```


## plot histogram for residuals distribution
```{r}
ggplot(bikemod_transform,aes(bikemod_transform$residuals))+
  geom_histogram()
```

## Prediction accuracy
```{r}
pred <- predict(bikemod_transform,newdata = testdata)
predictions <- pred^2
#MAE of the transformed regression model
print(mae(predictions,testdata$Bike)) 

#Plot predicted value vs. true values
par(mfrow=c(1,1))
plot(testdata$Bike,predictions,xlab=("True Bike Count"),ylab=("Predicted value"))
abline(1,1,col="red")
```

